---
title: "Finding high certainty levels for simple 1d model"
author: "Joel Eliason"
date: "March 20, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.width=12, fig.height=6)
library(deSolve)
library(BMS)
library(BAS)
library(gridExtra)
library(grid)
library(ggplot2)
source("simple1dsolve.R")
source("cubic2dsolve.R")
source("plot_cubic.R")
source("bms_summary.R")
set.seed(123)
```

Here, I will principally be concerned with finding initial conditions/data points that will give 99% certainty reconstruction of the simple 1d model and then paring data points/ICs back until we lose that level of certainty.

First, we have a uniform IC density (of 100 ICs) from 0 to 2, with 101 time steps per IC. This allows us to obtain transient behavior for each of these ICs, but not much steady state behavior.
```{r}
set.seed(123)
step=.01
tf=1
sd=0.1
degree=4
times=seq(0,tf,by=step)
x0=1

ic_step=0.02
ic_f=2
ics=seq(0.01,ic_f,by=ic_step)

df<-lapply(ics, function(z) simple1dsolve(x0=z,step=step,tf=tf,sd=sd,degree=degree))
df<-do.call("rbind",df)

fit<-bas.lm(y1~.,data=df)
summary(fit)
```

Next, let's try 100 ICs, but in a much smaller interval: 0 to 1e-4, with the same number of time steps.

```{r}
ics=seq(0.00001,0.0001,length.out=100)

df<-lapply(ics, function(z) simple1dsolve(x0=z,step=step,tf=tf,sd=sd,degree=degree))
df<-do.call("rbind",df)

fit<-bas.lm(y1~.,data=df)
summary(fit)
```

100 ICs, [1e-5,1e-4], tf=16
```{r}
step=.01
sd=0.1
degree=4

x0=1

ics=seq(0.00001,0.0001,length.out=100)
tf=16
times=seq(0,tf,by=step)

df<-lapply(ics, function(z) simple1dsolve(x0=z,step=step,tf=tf,sd=sd,degree=degree))
df<-do.call("rbind",df)

fit<-bas.lm(y1~.,data=df)
summary(fit)
```

At this range of ICs, with this number of ICs, tf=16 looks to be about as good as it's going to get (we can go out to about tf=40, but that only gives slight improvements in the post prob).

Let's try cutting our ICs down by half:

```{r}
step=.01
sd=0.1
degree=4

x0=1

ics=seq(0.00001,0.0001,length.out=50)
tf=16
times=seq(0,tf,by=step)

df<-lapply(ics, function(z) simple1dsolve(x0=z,step=step,tf=tf,sd=sd,degree=degree))
df<-do.call("rbind",df)

fit<-bas.lm(y1~.,data=df)
summary(fit)

```

Then to numICs=10:

```{r}
step=.01
sd=0.1
degree=4

x0=1

ics=seq(0.00001,0.0001,length.out=10)
tf=16
times=seq(0,tf,by=step)

df<-lapply(ics, function(z) simple1dsolve(x0=z,step=step,tf=tf,sd=sd,degree=degree))
df<-do.call("rbind",df)

fit<-bas.lm(y1~.,data=df)
summary(fit)

```

That's pretty significant. Let's go back to 50 and spread them out over a larger interval:

```{r}
step=.01
sd=0.1
degree=4

x0=1

ics=seq(0.00001,0.001,length.out=50)
tf=16
times=seq(0,tf,by=step)

df<-lapply(ics, function(z) simple1dsolve(x0=z,step=step,tf=tf,sd=sd,degree=degree))
df<-do.call("rbind",df)

fit<-bas.lm(y1~.,data=df)
summary(fit)

```

Let's spread them by two more orders of magnitude:

```{r}
step=.01
sd=0.1
degree=4

x0=1

ics=seq(0.00001,0.1,length.out=50)
tf=16
times=seq(0,tf,by=step)

df<-lapply(ics, function(z) simple1dsolve(x0=z,step=step,tf=tf,sd=sd,degree=degree))
df<-do.call("rbind",df)

fit<-bas.lm(y1~.,data=df)
summary(fit)

```

And two more:

```{r}
step=.01
sd=0.1
degree=4

x0=1

ics=seq(0.00001,10,length.out=50)
tf=16
times=seq(0,tf,by=step)

df<-lapply(ics, function(z) simple1dsolve(x0=z,step=step,tf=tf,sd=sd,degree=degree))
df<-do.call("rbind",df)

fit<-bas.lm(y1~.,data=df)
summary(fit)

```

Spreading it out over that range seemed like a good idea - let's now cut down the number of ICs again:
```{r}
step=.01
sd=0.1
degree=4

x0=1

ics=seq(0.00001,10,length.out=2)
tf=16
times=seq(0,tf,by=step)

df<-lapply(ics, function(z) simple1dsolve(x0=z,step=step,tf=tf,sd=sd,degree=degree))
df<-do.call("rbind",df)

fit<-bas.lm(y1~.,data=df)
summary(fit)

```

Let's try both ICs, independently:

```{r}
# set.seed(123)
step=.01
sd=0.1
degree=4

x0=1

ics=0.00001
tf=16
times=seq(0,tf,by=step)

df<-lapply(ics, function(z) simple1dsolve(x0=z,step=step,tf=tf,sd=sd,degree=degree))
df<-do.call("rbind",df)

fit<-bas.lm(y1~.,data=df)
summary(fit)

```

And IC=10:

```{r}
set.seed(123)
step=.01
sd=0.01
degree=4

x0=1

ics=10
tf=16
times=seq(0,tf,by=step)

df<-lapply(ics, function(z) simple1dsolve(x0=z,step=step,tf=tf,sd=sd,degree=degree))
df<-do.call("rbind",df)

fit<-bas.lm(y1~.,data=df)
summary(fit)

```

IC=10 is the best one, run for this long amount of time. Let's take a quick look at the solution for IC=10 again:

```{r}
plot(times,df[,2])
```

We're spending a lot of time in the same place. Let's cut down how long we need to sample by half:

```{r}
step=.01
sd=0.1
degree=4

x0=1

ics=10
tf=0.5
times=seq(0,tf,by=step)

df<-lapply(ics, function(z) simple1dsolve(x0=z,step=step,tf=tf,sd=sd,degree=degree))
df<-do.call("rbind",df)

fit<-bas.lm(y1~.,data=df)
summary(fit)

```
We definitely lose some probability mass, which is odd.

Let's fix IC=10, tf=3 and repeat these experiments over a lot of different random seeds.

```{r,warning=FALSE}
step=.01
sd=0.1
degree=4

ics=10
tf=0.5
times=seq(0,tf,by=step)

numSeeds=100
post_probs<-rep(0,numSeeds)
for (i in 1:numSeeds){
  set.seed(i)
  df<-simple1dsolve(x0=ics,step=step,tf=tf,sd=sd,degree=degree)
  fit<-bas.lm(y1~.,data=df)
  w<-fit$which
  lv=sapply(w,function(z) all(unlist(z)==c(0,1,3,4)))
  idx=min(which(lv == TRUE))
  post_probs[i]=fit$postprobs[idx]
}
hist(post_probs,breaks=40)
median(post_probs)
```